---
title: "Data 607 Assignment 9"
author: "Adam Gersowitz"
date: "3/27/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

In this assignment I will be using the New York Times API that I signed up for in order to download some of their information on best selling books into an r dataframe. I chose to look for fiction books that are currently on the New York Times fiction best seller list. First I found the appropriate name for this list which is "combined-print-fiction."

```{r}
library(httr)
library(jsonlite)
library(keyring)


#keyring::key_set(service = "my-database", 
#                 username = "mypassword")





api = keyring::key_get("my-database", "mypassword")

rn <- GET(paste("https://api.nytimes.com/svc/books/v3/lists/names.json?api-key=",api,sep=""))
rnbin<-(content(rn, "raw"))
writeBin(rnbin, "nytnames.txt")
nyt_names <- jsonlite::fromJSON("nytnames.txt")




url<-paste("https://api.nytimes.com/svc/books/v3/lists/combined-print-fiction.json?api-key=",api,sep="")

rf <- GET(url)
rfbin<-(content(rf, "raw"))
writeBin(rfbin, "nytfiction.txt")
nytf <- jsonlite::fromJSON("nytfiction.txt", simplifyDataFrame = TRUE)


nyt_fiction<-as.data.frame(nytf$results$books)
class(nyt_fiction)
#nyt_fiction



```

## Conclusion

I was successfully able to create a data frame base on the New York Times Combined Fiction list.Going forward I would want to iterate through more pages to get more results.





