---
title: "DATA 607 - Assignment 7"
author: "Adam Gersowitz"
date: "3/9/2020"
output:
  html_document: default
  pdf_document: default
---

## Overview

Assignment 7 will be focused on importing html,xml, and json tables into R and noting the differences between the 3 structures.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}


library(XML)
library(httr)
library(RCurl)
library(rvest)
library(rjson)
library(jsonlite)
library(tibble)
library(tidyr)
library(reshape)

xmlURL <- "https://raw.githubusercontent.com/agersowitz/Data-607-Datasets/master/books.xml"
xmldata = getURL(xmlURL)
booksxml<- xmlToDataFrame(xmldata)
booksxml<- data.frame(booksxml)


htmlURL <- "https://raw.githubusercontent.com/agersowitz/Data-607-Datasets/master/books.html"
htmlURL <- getURL(htmlURL)
#bookshtml <- htmlParse(xmldata)
bookshtml<- readHTMLTable(htmlURL,
                      stringsAsFactors = FALSE)
bookshtml<-data.frame(bookshtml)
#bookshtml_csv<-write.csv(bookshtml)


jsonURL <- "https://raw.githubusercontent.com/agersowitz/Data-607-Datasets/master/books.json"
jsondata = getURL(jsonURL)
jsonbooks <- fromJSON(jsondata)
booksjson<-data.frame(jsonbooks)

#json_raw <- enframe(unlist(jsonbooks))
#booksjson <- separate(json_raw, into = c(paste0("x", 1:10)))



#booksxml
#bookshtml
#booksjson

```

## Conclusions

I found that html, xml, and json all had unique challenges when loading them into a dataframe. I leaned heavily on Gaston Sanchez's slide deck as well as online resources like this tutorial (https://www.datacamp.com/community/tutorials/r-data-import-tutorial#data). I was able to create all three files and bring them into a data frame. This was a very good exercise for me,, because I have not used these formats extensivbely in the past so it is good to get practice with them due to their widespread use.

