---
title: "DATA 607 - Week 1 Assignment"
author: "Adam Gersowitz"
date: "2/8/2020"
output: html_document
---
## Overview

This is a basic example of connecting to a mysql database through R and then moving tables form that database into an R dataframe.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message = FALSE, warning = FALSE}

#install.packages('RMySQL')
#install.packages("RODBC")


library(RMySQL)
library(DBI)
library(dplyr)
library(dbplyr)
library(keyring)

#keyring::key_set(service = "my-database", 
#                 username = "mypassword")

#keyring::key_set(service = "my-database", 
#                 username = "myusername")
#dbDisconnect(con)
con <- dbConnect(RMySQL::MySQL(), 
                 user = keyring::key_get("my-database", "myusername"),
                 password=keyring::key_get("my-database", "mypassword"),
                 dbname = "flights")

#dbListTables(con)

final_dataset <- dbGetQuery(con,"
select reviews.review_id, reviews.rating, reviews.review, movies.*, people.*
from reviews join movies on movies.movie_id = reviews.movie_id 
join people on people.person_id = reviews.person_id
order by reviews.movie_id, reviews.person_id
")

reviews_df <- dbGetQuery(con,"select * from reviews order by review_id")
people_df <- dbGetQuery(con,"select * from people order by person_id")
movies_df <- dbGetQuery(con,"select * from movies order by movie_id")

write.csv(final_dataset,"C:/Users/gerso/Documents/DATA 607/Assignment 2/assignment 2 final dataset.csv", row.names = FALSE)
dbDisconnect(con)
```

## Conclusions

I broke this assignment out into 3 tables in my sql database. In my model the reviews table would only be populated when a review was made. This would help with data issues that would arise due to incomplete data, such as not every person has seen every movie. I would then reference the discriptor information of movies and people by joining the reviews table with a person_id or movie_id.





