---
title: "DATA 697 - Week 3 assignment"
author: "Adam Gersowitz"
date: "2/8/2020"
output:
  html_document: default
  pdf_document: default
---

## Overview

Week 3 assignment will be working with various data sets to practice Data manipulation and processing

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1 

Using the 173 majors listed in fivethirtyeight.com’s College Majors dataset [https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/], provide code that identifies the majors that contain either "DATA" or "STATISTICS"

```{r message = FALSE, warning = FALSE}
library (readr)
library(RCurl)


x <- getURL("https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/majors-list.csv")
majors <- read.csv(text=x)
#print(fightsongs)
#head(majors)

majorssdf <- data.frame(majors)
data_stats_majors <- subset(majorssdf, grepl("DATA", Major) | grepl("STATISTICS", Major))

data_stats_majors

```
## Question 2

Write code that transforms the data below:

[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"

[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  

[9] "elderberry"   "lime"         "lychee"       "mulberry"    

[13] "olive"        "salal berry"

Into a format like this:

c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")

```{r message = FALSE, warning = FALSE}
library(tidyverse)
list1 = c("bell pepper", "bilberry", "blackberry","blood orange")
list2 = c("blueberry","cantalope","chili pepper","cloudberry")
list3 = c("elderberry","lime","lychee","mulberry")
list4 = c("olive","salal berry")

final<-c(list1,list2,list3,list4)
final

x <- c("apple pie", "apple", "apple cake")

```

## Question 3

Describe, in words, what these expressions will match:

(.)\1\1

This will match any strings that have a character that repeats back to back

"(.)(.)\\2\\1"

This will match any strings that have a symetrical set of 4 characters, such that position 1 and 4 match while positions 2 and 4 match (ex: otto)

(..)\1

This will match any strings that have a repeated pair of letters

"(.).\\1.\\1"

This will match any string that has the same character repeat 3 times and they are all seperated by one character.
(ex: banana, papaya, x7xyx)

"(.)(.)(.).*\\3\\2\\1"

Similar to number 2 this will match any string that has a symetrical set of 6 characters. However this one will match is the first 3 characters are seperated from the last 3 characters. (ex:abccba, 123klsjd321) 



## Question 4

Construct regular expressions to match words that:

Start and end with the same character.

"^(.).*\\1$"


Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)

"(.)(.).*\\1\\2"


Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)

"(.).*\\1.*\\1"

```{r message = FALSE, warning = FALSE}
fruit <-c("banana","coconut","cucumber","jujubee","papaya","salal berry","elleven")

#str_view(fruit, "(.).*\\1.*\\1", match = TRUE)
#(.)\1\1
#"(.)(.)\\2\\1"
#(..)\1
#"(.).\\1.\\1"
#"(.)(.)(.).*\\3\\2\\1"

```